Forage site distributions show social and private information use in honeybees 

### 2021-11-29 analysis/manuscript/paper.Rmd
updated manuscript with new figures and vincent changes. Stil WIP before I can send it to Elli and Vincent

---
title: "Forage site distributions quantify collective foraging in honey bee colonies"
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

@@

Collective foraging requires that the emergent behaviour of an animal group takes a form that extends beyond the superposition of numerous individual contributions [@Sumpter2006], such that complexity is achieved at the level of the group or system, rather than the individual. Whilst collective foraging has been suggested to occur for many animal species (insert refs list), establishing if, and to what degree, animal groups make collective decisions remains a challenge. Here, we present a model of foraging by honey bees that can be used to distinguish collective decisions from individual-driven group behaviour. By confronting this model with waggle dance data from twenty hives, collected across two years, we can infer to what extent each colony depends on collective foraging vs individual decision-making. Our model both illustrates how a social insect group achieves superorganismal behaviour, and provides a tool to study variation in the occurrence of this behaviour across ecological scenarios.

# Main (Word count excluding figures and legend = 2357)

Collective decision making provides animals groups with the ability to pool individual experiences and make more informed decisions. Despite an enhanced capacity to process information, not every circumstance requires a collective response. In many cases, animal groups have been shown to do better when individuals prioritise information they obtain through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013]. Identifying the conditions which favour different information use strategies is therefore important, both for determining what conditions might have lead to the evolution of sociality and for understanding how animal groups might interact in an increasingly anthropomorphic world. However, assessing the importance of different information use strategies is not trivial and requires comparing the actions and behaviours of individuals with that of a collective. Often, information from different sources gets pooled together and so determining the importance of social Vs individual information use requires tools which can tease apart individual and collective behaviours.

@@

Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear. In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019]. In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008]. Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is logistically impossible due to honey bees small size, extensive foraging ranges and shear number of foragers. New methods which can quantify waggle dance use at a colony level are therefore needed if we are to assess the conditions where the dance provides the most benefit and derive the conditions user which this unique behaviour evolved.

Whether an individual forager follows a waggle dance from the dance floor or finds a resource for themselves, all foragers report resource information on the dance-floor [@Seeley1995]. Consequently, the difference in foraging behaviours between individuals following the dance and individuals relying on personal information should manifest itself in the pattern of dances reported on the dance-floor. To evaluate this idea, we simulated honey bees foraging in a landscape using both personal and waggle dance information (Fig. 1). As which individuals forage under what strategy is known in the simulation, we can compare the distributions of foraging distances reported on the dance floor. Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different. The distribution of the first simulation is akin to that of an exponential distribution (Fig. 1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods). The distribution of the distances reported by recruits (Fig. 1 D) is appears to resemble a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).

@@

Simulating honey bee foraging under a strategy of independent foraging and waggle dance following shows the shape of the foraging distance distributions reported on the dance-floor carries information about the extent to which honey bee colonies collectively make foraging decisions. On a real honey bee dance-floor, the distribution of distances is modified through the profitability of a resource reported. Honey bees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A]. Combined with the distance distributions identified from our simulations, we form a prediction of the distances reported on the dance-floor as a superposition of independent and waggle dance foraging (Fig 2, see Methods). This yields two different models: a collective model demonstrating collective decision making, comprised of the proportion of dances following an individual foraging strategy (scout) along with the proportion of dances following a waggle dance use strategy (recruit), and an individual model where decisions are made entirely individually (scout only) (Fig 2). By confronting these models with waggle dance data we can infer to what extent colony foraging depends on individual or collective decision making.

@@

To evaluate collective decision making in honey bee colonies foraging on 'natural' landscapes we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021]. Hives were situated at different locations in the Southeast of England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks. On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor. The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019]. For each site we fit both the collective and individual models. We used model selection to determine which model provided the better explanation of the data and model fit to determine if the model provided a good explanation of the data.

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a), however, in 3 sites the individual model provided the best explanation of foraging patterns. Across 17 of the sites, the collective model was not statistically different from the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites (Fig. 3.b). Sites ZSL (Fig. 3c) and STU (Fig 3d) are representative examples showing the model fits. For ZSL the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395). In contrast the scout model and empirical foraging distances are significantly different, providing a worse explanation of the data (AIC: 622, KS statistic: d = 0.122, p < 0.01). For STU the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).

@@

Our results indicate that, whilst foraging decisions are primarily reached collectively through the use of the waggle dance, in some circumstances, colony foraging is best explained by a model of individual foraging. This suggests that in some environments, honey bees actively ignore the waggle dance and instead prioritise the use of personal information. Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding feeders [@Gruter2011], suggesting experience drives information use at the individual level. Evaluating how these individual decisions influence the collective, however, has always been a challenge due to fact that it is effectively impossible to track an individuals foraging behaviour over a landscape.

As well as providing direct evidence that honey bees forage collectively according to a widely accepted theory [@Seeley1994; @Sherman2002], in showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors. These results support previous efforts to quantify collective foraging [@Sherman2002; @Dornhaus2006; @Gruter2013; @Price2019]. Furthermore, as our model predicts the proportion of dances coming from an individual or social foraging strategy, its fit to waggle dance data provides a tool to estimate a colonies use of the waggle dance.

To evaluate how waggle dance use is being influenced by the environment, we compared the different land-use types for each site to the estimated proportion of dances stemming from an individual foraging strategy (proportion of scout dances). Each site used in this study has a unique landscape comprised of different land-use types. We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments. As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the combination of land-use types which explained the most variation in the proportion of scout dances.

```{r, fig.cap="Figure 4. Estimated proportion of scouts for each site in the Agri-rural (left) and Urban (right) landscapes against the first principle component from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area."}
knitr::include_graphics("../results/figures/landuse_plots.png")
```

In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig. 4). This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space, built up areas and water. Arable land is typically nutritionally poor for honey bees (ref). On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref). Water also correlated negatively with the estimated proportion of scouts, possibly due to river and pond banks providing good floral resources for pollinators (ref). Combined, these results suggest when resources are sparse and difficult to find, as with arable land-use, individuals value personal information more than when easier to find, as with the non-agricultural unmanaged green space, built up areas and water.

```{r, agri-rual-loadings}
set_flextable_defaults(
  big.mark = " ",
  font.size = 10, theme_fun = theme_vanilla,
  padding.bottom = 6,
  padding.top = 6,
  padding.left = 6,
  padding.right = 6
)

load("../results/pls-loadings.RData")
argi.rural.loadings <- tibble(
  "Land-use" = names(agrirural_loadings$loadings),
  "PC1" = agrirural_loadings$loadings
)

urban.loadings <- tibble(
  "Land-use" = names(urban_loadings$loadings),
  "PC1" = urban_loadings$loadings
)

argi.rural.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
  flextable() %>%
  set_caption(
    "Land-use type loadings for the first principle component on the Agri-rural sites. Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
  ) %>%
  set_table_properties(layout = "autofit") %>%
  bold(
    i = ~ PC1 > agrirural_loadings$cutoff |
      PC1 < -agrirural_loadings$cutoff,
    j = "PC1"
  )
```

A similar effect is also observed in the urban environment, where a single PC explained ~61% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.61, $\phi$ = 7.4, p < 0.05, Fig. 4). This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland. This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit. As with the agri-rural land-uses, these results indicate when resources are sparse, as could be expected in continuous central and dense residential land, where there is limited space for floral resources, a colony emphasises the use of personal rather than social information. However, when resources are more abundant, as could be expected in sparse residential areas and amenity grasslands, a colony appears to increase the use of information contained in the waggle dance. Unlike in the agri-rural environment, water correlates positively with a higher use of the waggle dance. This likely indicates that the banks of urban water ways are less nutritionally rich than in agri-rural landscapes.

```{r, urban.loadings}
urban.loadings %>%
  mutate(PC1 = round(PC1, 3)) %>%
  mutate(`Land-use` = gsub("\\.", " ", `Land-use`)) %>%
  flextable() %>%
  set_caption(
    "Land-use type loadings for the first principle component on the urban sites. Bold face indicates loadings that contribute more than would be expected under a null hypothesis of equal contribution"
  ) %>%
  set_table_properties(layout = "autofit") %>%
  bold(
    i = ~ PC1 > urban_loadings$cutoff |
      PC1 < -urban_loadings$cutoff,
    j = "PC1"
  )
```

In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance. These results appear to follow what is predicted under theoretical models: in landscapes where resources are likely to more difficult to find, a colony tends to have a higher number of scouts than in landscapes where resources should be easier to locate [@Dornhaus2006; @Price2019]. However, it is important to note that other combinations which didn't show as having a significant association with the first principle component could have just as well fit into this argument, arguably better than others. In both the agri-rural and urban environments water seems to correlate in contrasting ways with the proportion of scouts. This could be due to differences in the floral richness of urban and agri-rural water ways (reff), however, if land-use type is to be taken as a proxy for floral availability, other variables such as parks, allotments and cemeteries in the urban environment and fruit or Oil Seed Rape in the agri-rural environment might be expected to be included in the first PC over water. An alternative explanation could be that water acts as barriers between land-use types and so gets bundled into the association (more thought on this needed I think).

Here, we demonstrate that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provide a methodology to further evaluate how honey bees use this unique behaviour. Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand .

Recent advances in computational methods have allowed individuals to be tracked within colonies in ways which would be almost impossible to do by hand, allowing theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021]. We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the model outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment. With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”. Our methodology provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.

>>>>

---
title: "Forage site distributions show social and private information use in honeybees "
author: "Joseph Palmer, Ash Samuelson, Elli Leadbeater and Vincent Jansen"
output:
  word_document
always_allow_html: true
bibliography: paper-refs.bib
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(showtext)
showtext_auto()

@@

Environmental conditions are widely understood to influence the value of social and private information when foraging. Exploring this relationship in detail requires quantifying the extent to which foraging decisions are made using social or private information. In honeybees, foragers use the waggle dance to organise collective foraging, yet under some conditions it’s use is less effective than prioritising private information. Identifying how information is used at a collective level is challenging due to the difficulty in evaluating individual behaviours on the colony scale. Here, we outline a model to quantify the extent to which honeybees use the waggle dance by evaluating the foraging distance distributions reported on the dance floor. In confronting this model with waggle dance data from twenty hives, we demonstrate that waggle dance use depends on the local environment of a hive. Further, by combining our estimates of waggle dance use with local landscape characteristics, our methods demonstrate a means of visualising the relationship between major landscape type and waggle dance use. These methods offer a novel approach to evaluate what factors likely led to the evolution of this unique behaviour, as well as helping determine how honeybees might adapt to an increasingly anthropogenic world.

# Main

Decision making by social insects typically requires the use of both social and private information (ref). However, to what degree social or private information is needed depends on circumstance (ref), and the mixture of private and social information varies both spatially and temporally. Although collective decision making using social information is an adaption to foraging in challenging environments (ref), in some cases, groups have been shown to do better when individuals prioritise information obtained through their own experiences to that provided by conspecifics [@King2007; @Rieucau2011; @Gruter2013]. Identifying the conditions which favour the use of these two types of information is important to understand how social insects make collective decisions, and determining what conditions might have led to the evolution of sociality as well as how social insects might respond to anthropogenic changes.

@@

Despite the efficiency benefits the waggle dance reportedly provides, its contribution to colony foraging success remains unclear. In many situations, colonies that are prevented from communicating via dance behaviour achieve equal or even greater foraging success than their wild-type counterparts [@Sherman2002; @Dornhaus2004; @Price2019]. In some ecological circumstances, individuals that forgo waiting for dance information in favour of individual search, termed “scouting”, may find food more quickly, and evidence suggests that bees adopt this strategy if the benefits of dance-following drop [@Gruter2008]. Whilst individuals can be observed under experimental conditions to determine how they use different sources of information, translating this up to a colony level is challenging due to honeybees small size, extensive foraging ranges and shear number of foragers. New methods which can quantify waggle dance use at a colony level are therefore needed if we are to assess the conditions where the dance provides the most benefit and derive the conditions user which this unique behaviour evolved.

Whether an individual forager uses social information and follows a waggle dance from the dance floor or acquires private information through scouting, all foragers report resource information on the dance-floor [@Seeley1995]. Consequently, the difference in foraging behaviours between recruited individuals following dances and scouting individuals relying on private information should manifest itself in the pattern of dances reported on the dance-floor. Here, we will demonstrate that in honeybees it is possible quantify the use of social and private information through the observation of honeybees’ waggle dances.

To reveal what foraging patterns might look like under a social or personal information use strategy, we simulated honeybees foraging in a landscape using both personal and waggle dance information (Fig. 1). As it is known in the simulation which individuals forage under what strategy, we can compare the distributions of foraging distances reported on the dance floor. Figs 1c and 1d show that the shapes of the resource distance distributions under the two simulated scenarios are very different. The distribution of the first simulation is akin to that of an exponential distribution (Fig. 1 C), which is the nearest neighbour distance distribution for foragers operating in a one-dimensional environment (see methods). The distribution of the distances reported by recruits (Fig. 1 D) appears to be a Rayleigh distribution which is the nearest neighbour distribution of a two-dimensional environment (see methods).

@@

Scouts and recruits report different foraging distance distributions on the dance-floor. This means the distances reported on the dance floor can be used to quantify the relative use of private and social information. Foragers that use private information (scouts) report a distribution of distances that reflects their probability of encountering a resource in the environment (see methods), whilst foragers that use social information (recruits) report distances that reflect the distribution of the most profitable resource available (see methods). The distance distribution reported on the dance floor therefore carries information about the extent to which private and social information is used.

On the dance-floor, the distribution of distances is modified through the profitability of a resource reported. Honey bees achieve this by measuring the energetic efficiency of a foraging trip through the ratio of energetic gain to energetic cost [@Seeley1994A]. Combined with the distance distributions identified in our simulations, we can describe the distances reported on the dance-floor as a superposition of private and social information use (Fig 2, see Methods). We assume that scouts report distances as they encounter resources on their scouting forays. If resources are distributed randomly this is approximated by an exponential distribution. Recruits report the most profitable resource among the resources advertised on the dance floor and follow a Rayleigh distribution (see Methods).

This yields two models for foraging: one model based on collective decision making which uses a mixture of private and social information, comprised of the proportion of dances following an individual foraging strategy (scouts), denoted by the parameter $p$, along with the proportion of dances, $1-p$, following a waggle dance use strategy (recruit), and a model based on the sole use of private information (scouts only, $p = 1$) (Fig 2). By using maximum likelihood methods and model selection, we inferred if and, by estimating the parameter $p$, to what extent colonies use social information.

@@

To evaluate collective decision making in honey bee colonies foraging on 'natural' landscapes we analysed a dataset of 3378 waggle dance observations from 20 observation hives between April-September 2017, previously described in [@Samuelson2021]. Hives were situated at different locations in South East England (see methods, figure 3A) and were visited every two weeks for a period of 24 weeks. On each visit, two hours of continuous waggle dance data were recorded by training a camcorder onto the dance floor. The footage of the dances was decoded manually to identify reported foraging locations [@Couvillon2012; @Schurch2019]. For each site we fit both the collective and individual foraging models. We used model selection to determine which model provided the better explanation of the data, and quantify the relative use of social information through estimating the parameter $p$. We calculated goodness-of-fit using a Kolmorgorov-Smirnov (KS) test to ascertain if the model provided a plausible explanation of the data.

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a), in 3 sites the individual model provided the best explanation of foraging patterns. Across 17 of the sites, the collective model had a good fit to the empirical foraging distances, whereas the scout model was significantly different to the data in 8 of the sites, and therefore not deemed a plausible model (Fig. 3.b). The sites shown in Figs  3c-d are representative examples showing the model fits. For the data in Fig 3c the collective model describes foraging distances extremely well and no significant difference is observed between the model and empirical foraging distances (AIC: 569, Bootstrapped KS statistic: d = 0.06, p = 0.395). In contrast the scout model and empirical foraging distances are significantly different, providing a worse explanation of the data (AIC: 622, KS statistic: d = 0.122, p < 0.01). For Fig 3d the collective model provides a slightly better fit (KS: d = 0.04, p = 0.999) than the scout model (KS: d = 0.05, p = 0.998), however, the scout model is more parsimonious (AIC: 388) than the collective model (AIC: 393).

@@

Our results show that, whilst foraging decisions are primarily reached using social and private information, in some circumstances, colony foraging is best explained by a model of individual foraging and private information use. This suggests that in some environments, honeybees actively ignore the waggle dance and instead prioritise the use of private information. Previous empirical studies have shown that individual foragers flexibly alter their use of waggle dance information in response to unrewarding feeders [@Gruter2011], suggesting experience drives information use at the individual level. Evaluating how these individual decisions influence the collective, however, has historically been a challenge as it is effectively impossible to track an individuals foraging behaviour over a landscape.

As well as providing direct evidence that honey bees forage collectively according to a widely accepted theory [@Seeley1994; @Sherman2002], in showing that foraging patterns in three of our sites are best explained by individual foraging, our results demonstrate honey bee colonies tune waggle dance use in relation to environmental factors.

Using the $p$ parameter estimated from our model fits, we evaluated how the use of social information varies across different environments. Each site used in this study has a unique landscape comprised of different proportions of common land-use types. We first classified the different land-use types of the area surrounding each site to obtain a standardised land-use profile for the urban and agri-rural environments separately. As many of the resulting land-use types covaried with each other, we performed a Partial Least Squares (PLS) analysis to determine the combination of land-use types which explained the most variation in the proportion of scout dances.

In the Agri-rural environment a single PC explained ~60% of the variation in the proportion of scouts (beta regression: $R^2$ = 0.6, $\phi$ = 4.9, p < 0.05, Fig. 4. A). This PC correlates positively with arable land and negatively with non-agricultural unmanaged green space (NAUMGS), built-up areas and water (Fig. 4. B). This analysis allows us to visualise what aspects of a colonies local environment most influence waggle dance use. As arable land increases whilst NAUMGS and built-up areas decrease, the proportion of individuals prioritising private information increases.

```{r, fig.cap="Figure 4. A. Estimated proportion of scouts for each site in the Agri-rural landscape against the first principle component derived from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area. B. Loadings for PC1 showing the correlations between PC1 and each land-use type (note NAUMGS stands for Non-agricultural unmanaged green space). All points outside the shaded area significantly contribute to PC1. The colours of these land-use types are shown in the local environment of sites HOR, SWP and CAD (maps in bottom left) chosen as representative sites along the model gradient."}
knitr::include_graphics("../results/figures/lu_prop_scouts_agrirural.png")
```

A similar effect is also observed in the urban environment, where a single PC explained ~61% of the variance in the proportion of scouts (beta regression: $R^2$ = 0.61, $\phi$ = 7.4, p < 0.05, Fig. 5. A). This PC correlates positively with continuous central land, dense residential land and water, whilst correlating negatively with sparse residential and amenity grassland (Fig. 5. B). This relationship appears to be largely driven by a single site showing a model of only individual foraging as the best fit. Visualising how waggle dance use is influenced by changes in these land-use types shows that when the landscape is dominated by sparse residential land, as in site HOR, waggle dance use dominates the colonies foraging strategy. However, this reduces as sparse residential land is replaced by continuous central and dense residential lands (Fig. 5)

```{r, fig.cap="Figure 4. A. Estimated proportion of scouts for each site in the urban landscape against the first principle component derived from a Partial Least Squares analysis of land-use type. A beta regression shows the relationship (black line) between PC1 and the proportion of scouts, with 95% CI shown by the blue shaded area. B. Loadings for PC1 showing the correlations between PC1 and each land-use type. All points outside the shaded area significantly contribute to PC1. The colours of these land-use types are shown in the local environment of sites MEL, SOM and BEL (maps in bottom left) chosen as representative sites along the model gradient."}
knitr::include_graphics("../results/figures/lu_prop_scouts_urban.png")
```

The fitting of our model to waggle dance observations provides an innovative new tool to quantify the relative importance of social and private information in honeybees. Further, by identifying the land-use combinations which most influence variation in the proportion of scouts, our analysis allows us to map the environment along the major axis of honeybee information use and visualise how land-use influences the use of the waggle dance. Despite being limited to 10 sites in an urban and agri-rural environment, our results appear to agree with previous findings that the environment influences waggle dance use [@Dornhaus2006; @Price2019].

In both the agri-rural and Urban environments, land-use types which associate with resource scarcity correlate positively with a higher proportion of scouts, whilst those which correlate with resource abundance correlate with a lower proportion of scouts, indicating a greater use of the waggle dance. Arable land is typically nutritionally poor for honey bees (ref). On the contrary, non-agricultural unmanaged green space is typically nutrient rich, as are built up areas in agri-rural environments, which often have gardens with ample flowers available for pollinators (ref). The same can also be said of the urban land-use types, where continuos central and dense residential areas are also nutritionally poor (ref) compared to sparse residential and amenity grasslands. In both landscapes, water shows as significant contributor but in opposite directions, increasing waggle dance use in the agri-rural environment whilst decreasing it in the urban. This might indicate that agri-rural water ways provide good floral resources for honeybees, whereas urban water ways do not. However, the map visualisations suggests that these land-use types offer little contribution to the overall environment. More data is therefore required to evaluate this relationship further.

Overall, we have demonstrated here that the analysis of waggle dance data can make an important contribution to our understanding of social information use and provides a methodology to further evaluate how honey bees use the unique waggle dance. Recently, technological advances have emerged which enable colony metrics to be collected faster, more accurately and over greater time spans than could be gathered by hand, allowing individuals to be tracked within colonies and theories of individual behaviour to be evaluated in more depth than could have been done previously [@Wild2021]. We are not yet in a position to be able to track behaviours outside the colony in the same detail, nevertheless, by piecing together the behavioural response of individuals into a mathematical description of colony foraging, the methodology outlined here extends our ability to predict individual behaviour beyond the colony walls and in to the environment. With the advances in the decoding of the waggle dance through automated methods [@Wario2017], we face the prospect of waggle dance data becoming “big data”. Our methodology thus provides a means of analysing such large data sets to inform the debate about the importance of social information, as well as providing useful colony metrics of foraging activity.
